Assessing Replication via P-values

### 2018-06-29 McNemar's Test/draft_mcnemars_ttest.Rmd
edits based off critics for t-test.

Unlike McNemar's test, the t-test assesses replication by directly comparing original and replication \textit{p}-values within a study pair. The null hypothesis for the t-test states that the original and replication studies have the same \textit{p}-value (i.e. $H_0: p_{1j} = p_{2j}$). The test statistic is given by $t = \frac{\bar{X}_{diff}}{SD_{diff}/\sqrt{k}}$ where $\bar{X}_{diff}$ is the mean of all differences between original and replication \textit{p}-values (i.e. $\bar{X}_{diff} = \frac{1}{k} \sum\limits_{j=1}^k (p_{1j} - p_{2j})$), $SD_{diff}$ is the standard deviation of those differences from the mean, and $k$ is the number of study pairs. A paired t-test is used in this case since the original and replication studies in each study pair are conducting the same experiment, using the same methods, and drawing conclusions on the same phenomenon.

@@

Like McNemar's test, the power of the t-test to detect if the \textit{p}-values of the original studies are different from those of the replication studies depends upon the power of both the original and replication studies. Consider a single study (either original or replication), with the null hypothesis $H_0:\theta_i = 0$. The finding from this study is significant if $\frac{|T_{i}|}{\sqrt{v_i}} \geq 1.96$ which is equivalent to $p_i \leq 0.05$ since $p_i = \Phi^{-1} \Big ( \frac{|T_i|}{\sqrt{v_i}} \Big)$ where $\Phi^{-1}$ is the standard normal quantile function.. Assuming that the variance is known for such a study and is $\frac{4}{n}$ where $n$ is the sample size then, the power to detect if $\theta_i$ is statistically different from 0 depends on the estimate $T_i$ and the sample size $n$. Extending this to the test statistic for the t-test, we note that the mean of the paired difference of \textit{p}-values can be written as

$$
\frac{1}{k}\sum\limits_{j=1}^k \Big [ \Phi^{-1} \Big ( \frac{|T_{1j}|}{\sqrt{v_{1j}}} \Big) - \Phi^{-1} \Big( \frac{|T_{2j}|}{\sqrt{v_{2j}}} \Big) \Big]
$$

>>>>

Unlike McNemar's test, the t-test assesses replication by directly comparing original and replication \textit{p}-values within a study pair. The null hypothesis for the t-test states that the original and replication studies' mean difference is zero (i.e. $H_0: \mu_{p_{1j}- p_{2j}}=0$). The test statistic is given by $t = \frac{\bar{X}_{diff}}{SD_{diff}/\sqrt{k}}$ where $\bar{X}_{diff}$ is the mean of all differences between original and replication \textit{p}-values (i.e. $\bar{X}_{diff} = \frac{1}{k} \sum\limits_{j=1}^k (p_{1j} - p_{2j})$), $SD_{diff}$ is the standard deviation of those differences from the mean, and $k$ is the number of study pairs. A paired t-test is used in this case since the original and replication studies in each study pair are conducting the same experiment, using the same methods, and drawing conclusions on the same phenomenon.

@@

Like McNemar's test, the power of the t-test to detect if the \textit{p}-values of the original studies are different from those of the replication studies depends upon the power of both the original and replication studies. Consider a single study (either original or replication), with the null hypothesis $H_0:\theta_i = 0$. The finding from this study is significant if $\frac{|T_{i}|}{\sqrt{v_i}} \geq 1.96$ which is equivalent to $p_i \leq 0.05$ since $p_i = \Phi^{-1} \Big ( \frac{|T_i|}{\sqrt{v_i}} \Big)$ where $\Phi^{-1}$ is the standard normal quantile function.. Assuming that the variance is known for such a study which is a function of the sample size $n_i$, then the power to detect if $\theta_i$ is statistically different from 0 depends on the estimate $T_i$ and the sample size $n_i$. Extending this to the test statistic for the t-test, we note that the mean of the paired difference of \textit{p}-values can be written as

$$
\mu_{p_{1j}- p_{2j}} = \frac{1}{k}\sum\limits_{j=1}^k \Big [ \Phi^{-1} \Big ( \frac{|T_{1j}|}{\sqrt{v_{1j}}} \Big) - \Phi^{-1} \Big( \frac{|T_{2j}|}{\sqrt{v_{2j}}} \Big) \Big]
$$
